Learning CNN filters from user-drawn image makers for coconut-tree image classification

### 2020-06-08 main.tex
Updates from Overleaf

\begin{abstract}
 Identifying species of trees in aerial images is essential for land-use classification, plantation monitoring, and impact assessment of natural disasters. The manual identification of trees in aerial images is tedious, costly, and error-prone, so automatic classification methods are necessary. Convolutional Neural Network (CNN) models have well succeeded in image classification applications from different domains. However, CNN models usually require intensive manual identification to create large training sets. We present a training method for CNN feature extraction that needs only a minimal set of images selected by the user. The method learns the filters of each convolutional layer from user-drawn markers in image regions that discriminate classes, allowing direct expert interference in the training process. It does not rely on optimization based on backpropagation, which considerably reduces the number of required images. We demonstrate its advantages on the binary classification of coconut-tree images against a popular CNN model.
\end{abstract}

\section{Introduction}
Deep learning has proven to be applicable to different tasks, from image classification to data synthesis \cite{goodfellow2016deep}. In remote sensing, the applications have involved segmentation of terrain images \cite{kemker2018algorithms, kampffmeyer2016semantic, hamaguchi2018effective}, building identification \cite{xu2018building, lu2018detecting, liu2018multilevel}, and deforestation monitoring \cite{bragilevsky2017deep}. In this work, we are interested in 
 identifying species of trees from aerial images. The topic is important for land-use classification, plantation monitoring during a season, and damage assessment of natural disasters, such as drilling and flooding. As these plantations can span very large areas, manual identification can be costly, tedious, and susceptible to errors, so automatic methods are necessary.

Classification of tree species in aerial images has been actively investigated~\cite{fassnacht2016review}. In \cite{puttemans2018comparing} and~\cite{aparna2018cnn}, the authors present automatic solutions to detect coconut trees based on convolutional neural network (CNN) models. Despite these recent advances, CNN models usually require considerable human effort in image annotation to create large training sets. Vargas-Muñoz et al.~\cite{8899005}  propose an active learning approach to reduce human effort in annotating images of coconut trees. They divide the problem in two stages: a stage that produces candidate image patches for having or not a coconut tree and a stage in which a CNN identifies the candidate patches with a coconut tree. We propose a different way to produce this classifier.

methods based on deep learning require a lot of annotated images for training, and annotating this volume of data is quite costly. Even interactive methods require large training sets. If these methods depend on huge training sets, their application may not be feasible in areas where it is impossible, or the cost of gathering and annotating these data sets is prohibitive. Much of the user's interaction with methods based on deep learning occurs through the annotation of the training set, and the user does not have much influence on the training process.

Despite the advances in deep learning over the last decades, especially with most recent deep neural networks, some important questions related to human interaction remain unanswered: (1) How to choose the most reasonable model, in the sense that it provides acceptable effectiveness, being as simple and efficient as needed, for a given classification problem? (2) How to train that model from a reduced number of samples?  (3) Can the user explain the decisions of that model? (4) Can that model improve from human corrections? The first question requires to explore human knowledge about the classification technique and the problem of interest. The second one requires to reduce as much as possible the human effort to train a model. The third issue is related to human understanding, and it may explore techniques (such as data visualization) to explain the decisions of the model and perhaps to guide human decisions concerning the design of the model. The fourth question is also important during training, and it is related to human control over the process. They all lead to the importance of involving human experts during the machine learning process.

In this paper, we present some advances towards the answers to the questions (1) and (2). We also discuss future research related to the questions (3) and (4). We present a method to the design of light-weighted convolutional neural networks from markers drawn by an expert on a few images to indicate representative regions of each object of interest for classification. From those markers, our approach can learn the filters of each layer, and the convolutional neural network can be designed in a layer-by-layer fashion, with no need for backpropagation in order to build a representative feature space for the classification problem. Backpropagation is only used to train a bank of multi-layer perceptrons for local pattern classification.

@@

\begin{figure}[!t]
  \centering
  \subfloat[\label{fig:ex-groups}]
  {\includegraphics[width=.35\linewidth]{figures/example_groups/groups_centered}}
  ~
  \subfloat[\label{fig:ex-groups-after}]{\includegraphics[width=.45\linewidth]{figures/example_groups/groups3d_centered}
  }
  \caption{Exemple of application of convolution with groups centers as filters' weightes: (a) Groups of three classes centered on the origin; (b) Groups after applying convolution and activation operations. }
  \label{fig:filter}
\end{figure}

@@

  \begin{tikzpicture}
    [
      grow                    = down,
      sibling distance        = 5em,
      level distance          = 3em,
      edge from parent/.style = {draw=black},
      every node/.style       = {treenode},
      sloped
    ]
    \node [treenode, fill=tealDeer, label = {[label distance = 0.5em]}] (v0) {Conv$(60 \times 7 \times 7)$}
      child {
        child {
          child {
                child {
                  node [treenode, fill=navyPurple] {Classifier}
                  edge from parent
                }
                node [treenode, fill=tealDeer] {BatchNorm}
                edge from parent
           }
          node [treenode, fill=tealDeer] {MaxPool$(3\times3)$}
          edge from parent
        }
        node [treenode, fill=dodgerBlue, label = {[label distance = 0.5em] }] {ReLU} 
        edge from parent
      };

@@

\begin{table}[!t]
  \begin{center}
  \begin{tabular}{|l|c|c|c|}
  \hline
   Method & Precision & Recall & F-score \\
  \hline\hline
    I-Ours & $\boldsymbol{0.86 \pm 0.01}$ & $\boldsymbol{0.84 \pm 0.02}$ & $\boldsymbol{0.85 \pm 0.02}$\\
    II-Ours (fine tuned) & $\boldsymbol{0.86 \pm 0.01}$ & $\boldsymbol{0.84 \pm 0.02}$ & $0.84 \pm 0.02$\\
    III-Ours with SVM & $0.836 \pm 0.015 $ & $ 0.793 \pm 0.025$ & $ 0.804 \pm 0.022$\\
    IV-VGG & $0.84 \pm 0.01$ & $0.76 \pm 0.03$ & $0.77 \pm 0.02 $ \\
    V-VGG (fine tuned) & $0.84 \pm 0.01$ & $0.75 \pm 0.03$ & $0.77 \pm 0.02 $ \\
  \hline
  \end{tabular}
  \end{center}
  \caption{Results.}
  \label{tab:results}
\end{table}

>>>>

\begin{abstract}
 Identifying species of trees in aerial images is essential for land-use classification, plantation monitoring, and impact assessment of natural disasters. The manual identification of trees in aerial images is tedious, costly, and error-prone, so automatic classification methods are necessary. Convolutional Neural Network (CNN) models have well succeeded in image classification applications from different domains. However, CNN models usually require intensive manual annotation to create large training sets. A CNN may be conceptually divided into convolutional layers for feature extraction and fully connected layers for classification. We present a method that needs a minimal set of user-selected images to train the CNN's feature extractor, considerably reducing the number of images to train its classifier. The method learns the filters of each convolutional layer from user-drawn markers in image regions that discriminate classes, allowing direct human expert interference in the training process. It does not rely on optimization based on backpropagation, which considerably reduces the number of required images. We demonstrate its advantages on binary classification of coconut-tree images against a popular CNN model.
\end{abstract}

\section{Introduction}
Deep learning has proven to be applicable to different tasks, from image classification to data synthesis \cite{goodfellow2016deep}. In remote sensing, the applications have involved segmentation of terrain images \cite{kemker2018algorithms, kampffmeyer2016semantic, hamaguchi2018effective}, building identification \cite{xu2018building, lu2018detecting, liu2018multilevel}, and deforestation monitoring \cite{bragilevsky2017deep}. In this work, we are interested in 
 identifying species of trees from aerial images. The topic is important for land-use classification, plantation monitoring, and damage assessment of natural disasters. As the plantations can span very large areas, the manual identification of each tree is costly, tedious, and error-prone, and so automatic classification methods are necessary.

Classification of tree species in aerial images has been actively investigated~\cite{fassnacht2016review}. In \cite{puttemans2018comparing} and~\cite{aparna2018cnn}, the authors present automatic solutions to detect coconut trees based on convolutional neural network (CNN) models. Despite these recent advances, CNN models usually require considerable human effort in image annotation to create large training sets. Vargas-Muñoz et al.~\cite{8899005} propose an approach to mitigate the problem based on active learning. The method explores data projection techniques to allow simultaneous annotation of multiple patches as having or not a coconut tree. It then uses a CNN to identify the candidate patches with a coconut tree. We adopt another alternative in this work -- the design of simplified CNN models from reduced annotated image sets and assistance of a human expert.

The design of CNN models without the expert as part of the training loop leaves several research questions unanswered: (1) How to choose the most reasonable model, in the sense that it provides acceptable effectiveness, being as simple and efficient as needed, for a given classification problem? (2) How to train that model from a reduced number of samples?  (3) Can the user explain the decisions of that model? (4) Can that model improve from human corrections? The first question requires to explore human knowledge about the classification technique and the problem of interest. The second one requires to reduce as much as possible the human effort to train a model. The third issue is related to human understanding, and it may explore data visualization techniques to explain the decisions of the model and perhaps to guide human decisions concerning the design of the model. The fourth question is also important during training, and it is related to human control over the process. They all lead to the importance of involving human experts during the machine learning process.

In this paper, we present some advances towards answering the questions (1) and (2). We also discuss future research related to the questions (3) and (4). We present a method to the design of simplified convolutional neural networks from markers drawn by a human expert on a few images to indicate the most discriminant regions among the classes. From those markers, our approach can learn the filters of each convolutional layer, such that the CNN can be created in a layer-by-layer fashion, with no need for backpropagation in order to build a representative feature space for the classification problem. Backpropagation is only used to train a bank of multi-layer perceptrons for local pattern classification.

@@

\begin{figure}[!t]
  \centering
  \subfloat[\label{fig:ex-groups}]
  {\includegraphics[width=.35\linewidth]{figures/example_groups/groups_centered}}
  ~
  \subfloat[\label{fig:ex-groups-after}]{\includegraphics[width=.45\linewidth]{figures/example_groups/groups3d_centered}
  }
  \caption{Example of application of convolution with groups centers as filters' weights: (a) Groups of three classes centered on the origin; (b) Groups after applying convolution and activation operations. }
  \label{fig:filter}
\end{figure}

@@

  \begin{tikzpicture}
    [
      grow                    = down,
      sibling distance        = 5em,
      level distance          = 3em,
      edge from parent/.style = {draw=black},
      every node/.style       = {treenode},
      sloped
    ]
    \node [treenode, fill=tealDeer, label = {[label distance = 0.5em]}] (v0) {Conv$(60 \times 7 \times 7)$}
      child {
        child {
          child {
                child {
                  node [treenode, fill=navyPurple] {Classifier}
                  edge from parent
                }
                node [treenode, fill=tealDeer] {BatchNorm}
                edge from parent
           }
          node [treenode, fill=tealDeer] {MaxPool$(3\times3)$}
          edge from parent
        }
        node [treenode, fill=tealDeer, label = {[label distance = 0.5em] }] {ReLU} 
        edge from parent
      };

@@

\begin{table}[!t]
  \begin{center}
  \begin{tabular}{|l|c|c|c|}
  \hline
   Method & Precision & Recall & F-score \\
  \hline\hline
    I-Ours & $0.856 \pm 0.014$ & $0.836 \pm 0.020$ & $0.842 \pm 0.018$\\
    II-Ours (fine tuned) & $0.86 \pm 0.01$ & $0.84 \pm 0.02$ & $0.84 \pm 0.02$\\
    III-Ours with SVM & $0.836 \pm 0.015 $ & $ 0.793 \pm 0.025$ & $ 0.804 \pm 0.022$\\
    IV-VGG & $0.84 \pm 0.01$ & $0.76 \pm 0.03$ & $0.77 \pm 0.02 $ \\
    V-VGG (fine tuned) & $0.84 \pm 0.01$ & $0.75 \pm 0.03$ & $0.77 \pm 0.02 $ \\
  \hline
  \end{tabular}
  \end{center}
  \caption{Results.}
  \label{tab:results}
\end{table}
